---
title: "Homework 4: Biggish Data dottable"
author: "Sloane Johns"
date: "2025-03-04"
output:
  pdf_document: default
  pdf: default
---

```{r, message=FALSE}
# libraries
library(tidyverse)
library(dplyr)
library(data.table)

# data
nycdata <- fread("~/hw_dottable/nycdata.csv")
```

#### Question 1:

Use and show data.table code to select the variables year, month, day, and hour from the imported flights data.
```{r}
head(nycdata[, .(year, month, day, hour)], 6)
```

#### Question 2:

Use and show data.table code to produce a table that shows a carrier of DL, and origin of JFK and a destination of SEA.
```{r}
head(nycdata[carrier == "DL" & origin == "JFK" & dest == "SEA"], 6)
```

#### Question 3:

Use and show data.table code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.
```{r}
head(nycdata[carrier == "UA" & month == 3 & air_time < 330], 6)
```

#### Question 4:

Use and show tidyverse code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.
```{r}
nycdata %>% filter(carrier == "UA" & month == 3 & air_time < 330) %>% head(6)
```

#### Question 5:

Use the data.table method to add a variable called speed that is the average air speed of the plane in miles per hour.
```{r}
nycdata[, c("speed") := .(distance / (air_time / 60))]
head(nycdata, 6)
```

#### Question 6:

Use the tidyverse method to add a variable called speed that is the average air speed of the plane in miles per hour.
```{r}
nycdata %>% mutate(speed = distance / (air_time / 60)) %>% head(6)
```

